Synthesis of Rules for Network Intrusion Detectors

### 2019-10-24 paper/paper.tex
updating background and technique

Suricata supports a total of \totoptions\ options covering
\numproto\ distinct protocols. Of these \totoptions,
\totoptionsrelevant\ options are relevant to capture traffic
\begin{wrapfigure}[17]{r}{0.52\textwidth}
  \centering
  \vspace{-4ex}
  \scalebox{.75}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
%          ylabel={\#rules},
          symbolic x coords={gen, http, icmp, udp, tcp},
          xtick=data,
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
        \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
        \legend{relevant, irrelevant}
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-rules-protocol}Distribution of
    number of options per protocol and relevance for triggering action.}
\end{wrapfigure}
Options can be specific
to a given protocol or general and they can influence or not
triggering an action. The purpose of rule option \CodeIn{msg}, for
instance, is to print on the output an informative message indicating
what kind of intrusion has been
observed. Figure~\ref{fig:distribution-rules-protocol} shows the
distributions of these options per protocol and their relevance for
triggering rule actions. We obtained this information
by\Fix{...explicar como...} \Fix{eh um pouco suspeito ter tantas
  regras irrelevantes...}

%% This plot indicates that the problem search
%% space is defined based on the protocol of the malicious traffic. For
%% instance, if \tname{} detects by analyzing the format of network
%% packets that the attack involves \CodeIn{http} traffic, then 34
%% different rule options will be considered in the creation of the rule.

Table~\ref{table:rules}
shows a sample of the relevant options of Suricata. Column ``\#''
shows the option number, column ``Name'' shows the name of the
option, column ``\#~Params'' shows the number of additional parameters
associated with a given option, and column ``Description'' presents a
short description of the option purpose. \Fix{...explain content...}

@@

\Fix{atualize isto para o valor correto, por favor. $\rightarrow$}
\begin{figure}[h!]
  \vspace{-3ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold}]{synflood.suricata.synth}
%  \caption{Suricata rule for SYN Flood Attacks.}
  %  \label{fig:synflood-example-synt}
  \vspace{-3ex}  
\end{figure}

>>>>

Suricata supports a total of \totoptions\ distinct options covering
\numproto\ distinct protocols. A total of 9 of these options are for
documentation.
\begin{wrapfigure}[16]{r}{0.5\textwidth}
  \centering
  \vspace{-5ex}
  \scalebox{0.7}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={dns, ftp, http, icmp, ip, smb, smtp, ssh,
          tcp, tls, udp},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(dns,10) (ftp,11) (http,20) (icmp,18)(ip,16) (smb,13) (smtp,11) (ssh, 10) (tcp, 32) (tls,18) (udp, 18)};
%        \legend{relevant, irrelevant}
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-rules-protocol}Number of relevant options per protocol.}
\end{wrapfigure}
The purpose of rule option \CodeIn{msg}, for instance, is to print on
the output an informative message indicating what kind of intrusion
has been observed. Figure~\ref{fig:distribution-rules-protocol} shows
the distributions of these options per protocol. Note that the sum of
the number of protocol options is higher than \totoptions\ as some
options can be used by multiple protocols. Table~\ref{table:rules}
shows a sample of the options of Suricata. Column ``\#'' shows the
option number, column ``Name'' shows the name of the option, column
``\#~Params'' shows the number of additional parameters associated
with a given option, and column ``Description'' presents a short
description of the option purpose. \Fix{...explain content...}

@@

\Fix{atualize isto para o valor correto, por favor. $\rightarrow$}

\begin{figure}[h!]
  \vspace{-3ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold}]{synflood.suricata.synth}
%  \caption{Suricata rule for SYN Flood Attacks.}
  %  \label{fig:synflood-example-synt}
  \vspace{-3ex}  
\end{figure}
